---
title: "FA2"
author: "Borromeo_Mayo_Mercado"
date: "2025-09-10"
output: 
  pdf_document:
---

**Github link: https://github.com/chieelo/STATS.git **

## 3.49 Prove

$$
\sum_{i=1}^n (x_i - 1)^2 = \sum_{i=1}^n x_i^2 - 2\sum_{i=1}^n x_i + n.
$$

### Proof

We know that

$$
(a - b)^2 = a^2 - 2ab + b^2.
$$

So, lets expand the left hand side of the equation.

$$
(X_j - 1)^2 = X_j^2 - 2X_j + 1.
$$

Now, sum over all $j = 1, 2, \dots, N$:

$$
\sum_{j=1}^N (X_j - 1)^2 = \sum_{j=1}^N \big(X_j^2 - 2X_j + 1\big).
$$

Then, we distribute the summation

$$
\sum_{j=1}^N (X_j - 1)^2
= \sum_{j=1}^N X_j^2 \;-\; 2\sum_{j=1}^N X_j \;+\; \sum_{j=1}^N 1.
$$

The third term is simply $N$, because we are adding $1$ exactly $N$ times

So:

$$
\sum_{j=1}^N (X_j - 1)^2 = \sum_{j=1}^N X_j^2 - 2\sum_{j=1}^N X_j + N.
$$

------------------------------------------------------------------------

Therefore, the right-hand side is just the expansion of the left-hand side.

$$
\sum_{j=1}^N (X_j - 1)^2 = \sum_{j=1}^N X_j^2 - 2\sum_{j=1}^N X_j + N. \blacksquare
$$

## 3.51

Two variables, U and V, assume the values:

U1 = 3, U2 = 2, U3 = 5

V1 = 4, V2 = 1, V3 = 6

# Calculate:
```{r}
# Define the variables
U <- c(3, -2, 5)
V <- c(-4, -1, 6)


a <- sum(U * V)


b <- sum((U + 3) * (V - 4))


c <- sum(V^2)


d <- sum(U) * (sum(V))^2


e <- sum(U * V^2)


f <- sum(U^2 - 2*V^2 + 2)


g <- sum(U / V)


```

\begin{flushleft}
\(\displaystyle (a)\; \sum U V \)
\end{flushleft}
 = **`r paste(a)`**

\begin{flushleft}
\(\displaystyle (b)\; \sum (U+3)(V-4) \)
\end{flushleft}
 = **`r paste(b)`**

\begin{flushleft}
\(\displaystyle (c)\; \sum V^2 \)
\end{flushleft}
 = **`r paste(c)`**

\begin{flushleft}
\(\displaystyle (d)\; (\sum U)(\sum V)^2 \)
\end{flushleft}
 = **`r paste(d)`**

\begin{flushleft}
\(\displaystyle (e)\; \sum U V^2 \)
\end{flushleft}
 = **`r paste(e)`**

\begin{flushleft}
\(\displaystyle (f)\; \sum (U^2 - 2V^2 + 2) \)
\end{flushleft}
 = **`r paste(f)`**

\begin{flushleft}
\(\displaystyle (g)\; \sum (U/V) \)
\end{flushleft}
 = **`r round(g, 2)`**




## 3.9

**Find the geometric mean of the sets:**

**(a)** $[3, 5, 8, 3, 7, 2]$

**(b)** $[28.5, 73.6, 47.2, 31.5, 64.8]$

**The geometric mean of a set of** $n$ numbers $x_1, x_2, \ldots, x_n$ is given by:

$$
GM = \left(\prod_{i=1}^n x_i \right)^{\frac{1}{n}}
$$



```{r setup}
a <- c(3, 5, 8, 3, 7, 2)
geom_mean_a <- prod(a)^(1/length(a))
```
Set A: $[3, 5, 8, 3, 7, 2]$   
Therefore, the geometric mean of set (a) is: **4.140681**



```{r}
b <- c(28.5, 73.6, 47.2, 31.5, 64.8)
geom_mean_b <- prod(b)^(1/length(b))
```
Set B: $[28.5, 73.6, 47.2, 31.5, 64.8]$  
Therefore, the geometric mean of set (b) is: **45.8258**