---
title: "FA8 Assumption 1 and 2"
author: "Borromeo Mayo Mercado"
date: "2025-11-10"
output:
  pdf_document: default
  html_document: default
---

---
title: "Independent Samples t-test: Invisibility Cloak Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following table is loaded from the csv file found in FA8 Instructions.

```{r load-data}
library(knitr)

df <- read.csv("InvisibilityCloak.csv", header = TRUE)
kable(df, caption = "Invisibility Cloak Dataset")

```

---

# Assumption 1: Dependent Variable Must Be Continuous

The dependent variable Mischief represents the number of mischievous behaviors.  
Even though Mischief is technically count data, the range (0–8) and multiple numeric values allow us to treat it as continuous for a t-test.

**Assumption 1 is satisfied.**

```{r assume1-table, warning=FALSE, message=FALSE}
suppressMessages(library(dplyr))
suppressMessages(library(knitr))

mischief_summary <- df %>%
  summarise(
    Min = min(Mischief),
    `1st Quartile` = quantile(Mischief, 0.25),
    Median = median(Mischief),
    Mean = mean(Mischief),
    `3rd Quartile` = quantile(Mischief, 0.75),
    Max = max(Mischief),
    SD = sd(Mischief)
  )

kable(mischief_summary, caption = "Descriptive Statistics for Mischief")
```

---

# Assumption 2: Independent Variable Must Have Two Independent Groups

The independent variable Cloak has exactly two groups:

0 = Without Cloak  
1 = With Cloak  

Additionally, there are equal number of participants in each group (12 participants each) and no participant belongs to both groups.

**Assumption 2 is satisfied.**

```{r assume2-table, warning=FALSE, message=FALSE}
suppressMessages(library(dplyr))
suppressMessages(library(knitr))

cloak_summary <- df %>%
  group_by(Cloak) %>%
  summarise(`Number of Participants` = n()) %>%
  mutate(Group = ifelse(Cloak == 0, "Without Cloak", "With Cloak")) %>%
  select(Group, `Number of Participants`)

kable(cloak_summary, caption = "Number of Participants per Cloak Group")
```








# Computation

## 1. Levene's Test for Equality of Variances

```{r levene-test, warning=FALSE, message=FALSE}
library(lawstat)
levene_result <- levene.test(df$Mischief, df$Cloak)

levene_table <- data.frame(
  "Test Statistic" = round(levene_result$statistic, 3),
  "p-value" = round(levene_result$p.value, 3)
)

knitr::kable(levene_table, caption = "Levene's Test for Equality of Variances")
```

This table shows whether the variances of Mischief are equal between the two groups.Since p = 0.609 > 0.05 then variances are not significantly different, so we can assume equal variances for the t-test.

---

## 2. Independent Samples t-test

```{r t-test, warning=FALSE, message=FALSE}
t_result <- t.test(Mischief ~ factor(Cloak), data = df, var.equal = TRUE)

group_means <- t_result$estimate
mean_diff <- diff(group_means)
ci_lower <- t_result$conf.int[1]
ci_upper <- t_result$conf.int[2]
t_value <- t_result$statistic
df_val <- t_result$parameter
p_val <- t_result$p.value

t_table <- data.frame(
  "Group" = c("Without Cloak", "With Cloak"),
  "Mean" = round(group_means, 2)
)

t_summary <- data.frame(
  "t-value" = round(t_value, 3),
  "df" = df_val,
  "p-value" = round(p_val, 3),
  "Mean Difference" = round(mean_diff, 3),
  "95% CI Lower" = round(ci_lower, 3),
  "95% CI Upper" = round(ci_upper, 3)
)

knitr::kable(t_table, caption = "Group Means for Mischief")
knitr::kable(t_summary, caption = "Independent Samples t-test Results")
```

These tables clearly summarize the group means and t-test results, including mean difference, confidence interval, and p-value. The t-test shows no significant difference between groups (t(22) = -1.713, p = 0.101), with a mean difference of 1.25 and 95% CI [-2.763, 0.263].






