---
title: "Summative Assessment 1"
author: "Borromeo_Mayo_Mercado"
date: "2025-10-08"
output:
  pdf_document:
    latex_engine: xelatex
---

# Part 1

A study was undertaken to compare the mean time spent on cell phones by male and female college students per week. Fifty male and fifty female students were selected from Midwestern University, and the number of hours per week spent talking on their cell phones was determined. The results in hours are shown in **Table 10.6**. It is desired to test: $$
H_0: \mu_1 = \mu_2 \quad \text{versus} \quad H_1: \mu_1 \neq \mu_2
$$

### Table 10.6 – Hours Spent Talking on Cell Phones

| **Males** |     |     |     |     |     | **Females** |     |     |     |     |
|:---------:|:---:|:---:|:---:|:---:|:---:|:-----------:|:---:|:---:|:---:|:---:|
|    12     |  4  | 11  | 13  | 11  |     |     11      |  9  |  7  | 10  |  9  |
|     7     |  9  | 10  | 10  |  7  |     |     10      | 10  |  7  |  9  | 10  |
|     7     | 12  |  6  |  9  | 15  |     |     11      |  8  |  9  |  6  | 11  |
|    10     | 11  | 12  |  7  |  8  |     |     10      |  7  |  9  | 12  | 14  |
|     8     |  9  | 11  | 10  |  9  |     |     11      | 12  | 12  |  8  | 12  |
|    10     |  9  |  9  |  7  |  9  |     |     12      |  9  | 10  | 11  |  7  |
|    11     |  7  | 10  | 10  | 11  |     |     12      |  7  |  9  |  8  | 11  |
|     9     | 12  | 12  |  8  | 13  |     |     10      |  8  | 13  |  8  | 10  |
|     9     | 10  |  8  | 11  | 10  |     |      9      |  9  |  9  | 11  |  9  |
|    13     | 13  |  9  | 10  | 13  |     |      9      |  8  |  9  | 12  | 11  |

```{r}
# install.packages("psych")
knitr::opts_chunk$set(echo = TRUE)
library(psych)
library(knitr)
male <- c(
  12, 4, 11, 13, 11,
  7, 9, 10, 10, 7,
  7, 12, 6, 9, 15,
  10, 11, 12, 7, 8,
  8, 9, 11, 10, 9,
  10, 9, 9, 7, 9,
  11, 7, 10, 10, 11,
  9, 12, 12, 8, 13,
  9, 10, 8, 11, 10,
  13, 13, 9, 10, 13
)

female <- c(
  11, 9, 7, 10, 9,
  10, 10, 7, 9, 10,
  11, 8, 9, 6, 11,
  10, 7, 9, 12, 14,
  11, 12, 12, 8, 12,
  12, 9, 10, 11, 7,
  12, 7, 9, 8, 11,
  10, 8, 13, 8, 10,
  9, 9, 9, 11, 9,
  9, 8, 9, 12, 11
)


male_female <- c(male, female)
```

------------------------------------------------------------------------

## Question 1: Provide descriptive statistical summaries of the entire data.

```{r}
description <- describe(male_female)
kable(description[, c("n", "mean", "sd", "min", "max", "skew", "kurtosis")],
      caption = "Table 1: Descriptive Statistics for All Students",
      digits = 2, align = "c")

```

**Interpretation:** Both male and female students, they spent almost 9.8 hours per week talking on their cellphones.

------------------------------------------------------------------------

## Question 2: Provide descriptive statistical summaries of the data for each gender category.

```{r}
description_m <- describe(male)
description_f <- describe(female)

summary_table <- data.frame(
  Statistic = c("N", "Mean", "Median", "SD", "Min", "Max", "Range"),
  Male = c(description_m$n,
            round(description_m$mean, 2),
            description_m$median,
            round(description_m$sd, 2),
            description_m$min,
            description_m$max,
            round(description_m$range, 2)),
  Female = c(description_f$n,
              round(description_f$mean, 2),
              description_f$median,
              round(description_f$sd, 2),
              description_f$min,
              description_f$max,
              round(description_f$range, 2))
)

kable(summary_table,
      caption = "Table 2: Summary Statistics by Gender",
      align = "c")


```

**Interpretation:** Male: Mean = 9.82 hours, SD = 2.05 Female: Mean = 9.50 hours, SD = 1.87 Men and women have almost the same number of hours.

------------------------------------------------------------------------

## Question 3: Make a report based on the statistical summaries, including the results for both the combined data (not split by gender) and the gender-specific categories.

```{r}
result<- t.test(male, female, alternative = "two.sided", var.equal = TRUE)

results_table <- data.frame(
  Statistic = c("t-value", "Degrees of Freedom", "p-value"),
  Result = c(round(result$statistic, 2),
             round(result$parameter, 0),
             round(result$p.value, 3))
)

kable(results_table, caption = "Table 3: Two-Sample t-test Results", align = "c")

```

**Interpretation:**

If $p < 0.05$, we **reject** $H_0$ and conclude that there is a **significant difference** in the mean hours males and females spend talking on cell phones.\
If $p > 0.05$, we **fail to reject** $H_0$ and conclude that there is **no significant difference**.

Based on the computed results:

$$
t ≈ 1.11, p ≈ 0.27
$$

Since $p > 0.05$, we **fail to reject the null hypothesis**.

***Therefore, we have no strong evidence that there is a significant difference between genders in their weekly phone usage.***
